---
output:
  html_document: 
    theme: cerulean
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r setKnitrOptions, include=FALSE}
knitr::opts_chunk$set(results='asis', echo = FALSE, error = TRUE, cache = FALSE, cache.comments = TRUE, fig.dim = c(7,7), autodep = TRUE)
```




```{r loadParamsAndDependencies, warning=FALSE, message=FALSE}
### Define working folder (contains R/Rmd file for current sample, parent contains global project files)
WORKING_DIR   = getwd();

### Load parameters
# Define an environment that will contain parameters
paramsEnv = new.env();

# Load file defining global (project) parameters in 'paramsEnv'
source( file.path( WORKING_DIR, "..", "globalParams.R"), local = paramsEnv);

# Load file defining local (sample) parameters eventually overriding global ones
sampleParamsFilePath = file.path( WORKING_DIR, "..", "sampleParams.R");
if(file.exists(sampleParamsFilePath)){
  source( sampleParamsFilePath, local = paramsEnv);
} else {
  warning("The file 'sampleParams.R' containing sample-specific parameters is missing. The analysis will use global (project) parameters only.");
}

# Load file defining analysis parameters
analysisParamsFilePath = file.path( WORKING_DIR, "analysisParams.R");
if(file.exists(analysisParamsFilePath)) {
  source( analysisParamsFilePath, local = paramsEnv);
} else {
  warning("The file 'analysisParamsFilePath.R' containing analysis-specific parameters is missing.");
}

# Assign loaded values to current environment (Not .GlobalEnv as it may not be the current one depending on where rendering is started from)
invisible( lapply( ls( paramsEnv, all.names = TRUE), function(x, envir)
  { 
    assign( x = x, 
            value = get( x, pos = paramsEnv), 
            pos = envir)
  }, 
  environment()));

### Load general Deps
cat("<BR>WORKING_DIR=", WORKING_DIR)
source( file.path( WORKING_DIR, "00_generalDeps.R"));

# Import chunks from source files
read_chunk( path=file.path( WORKING_DIR, "01_prepareData.R"));
read_chunk( path=file.path( WORKING_DIR, "02_variableGenes.R"));
read_chunk( path=file.path( WORKING_DIR, "03_dataIntegration.R"));
read_chunk( path=file.path( WORKING_DIR, "04_macrophageSelection.R"));
```

---
title: `r paste( GLOBAL_DESCRIPTION, ANALYSIS_STEP_NAME, sep ="<BR>")`
---


```{r initFunctions}

### Create output directory
dir.create( PATH_ANALYSIS_OUTPUT, showWarnings = FALSE);


### Record starting time (for report and file names)
startTime = Sys.time();
startTimeReport   = format( startTime, "%Y %m %d %H:%M:%S")
startTimeFileName = format( startTime, "%Y_%m_%d_%H_%M_%S")


### Initialize utils

# Create a hash for all parameters
paramsHash = digest( paramsEnv, algo="murmur32");

# Create a 'safe' unique prefix for output files
# outputFilesPrefix = paste0( SCIENTIFIC_PROJECT_NAME, "_",     
#                             EXPERIMENT_PROJECT_NAME, "_"
#                             #startTimeFileName, "_",
#                             #paramsHash, "_"
#                             )

# Set seed for pseudo random number generator
set.seed(42);

# Create a hidden empty heatmap to prevent heatmapr interference with plotly functions (bug ?)
cat('<div style="display: none">');
htmltools::tagList(main_heatmap(matrix(0)));
cat('</div>');

# Create a hidden plotly to prevent bug when rendering multiple plots in report
cat('<div style="display: none">');
htmltools::tagList(plot_ly(x = rnorm(10), type = "histogram"));
cat('</div>');

# Create a hidden empty heatmap to prevent iheatmap interference with plotly functions
cat('<div style="display: none">');
htmltools::tagList(iheatmap(rnorm(10)));
cat('</div>');
```




## Primary analysis (QC)


### Run informations {.tabset .tabset-fade}

#### Run
<center>
```{r runInfo_run}
pander( c( 'Date'     = startTimeReport,
           'Hash'     = paramsHash));
```
</center>

#### Parameters
```{r runInfo_params}
#showSimpleDT( paramsEnv); # Does not handle lists or nested lists easily
listParams = as.list( paramsEnv);
pander(listParams[order(names(listParams))]); # Show params in alphabetical order
```

#### System
<center>
```{r runInfo_sysInfo}
showSimpleDT( Sys.info())
```
</center>

#### R
```{r runInfo_r}
pander( sessionInfo());
```

<!-- Ending the tabset section -->
### {.toc-ignore} 
<br /><br />

### Load data from quality control step
```{r dataLoadFilterNormalize}
<<loadData>>
```
<br /><br />


### Highly variable genes
#### Juvenile sample
```{r findHVG_rep1}
<<findVariableGenes_seuratMethodData_rep1>>
<<findVariableGenes_summaryPlot_rep1>>
<<findVariableGenes_summaryTable_rep1>>
```

#### Adult sample
```{r findHVG_rep2}
<<findVariableGenes_seuratMethodData_rep2>>
<<findVariableGenes_summaryPlot_rep2>>
<<findVariableGenes_summaryTable_rep2>>
```

#### Aged sample
```{r findHVG_rep3}
<<findVariableGenes_seuratMethodData_rep3>>
<<findVariableGenes_summaryPlot_rep3>>
<<findVariableGenes_summaryTable_rep3>>
```
<br /><br />


### Data integration {.tabset .tabset-fade}
This part shows the results of the combination of the data from the above replicate.
```{r dataIntegrationProcessing}
<<dataIntegration>>
<<analysisIntegration>>
```


#### tSNE
```{r visualizationIntegration_tSNE}
useReduction='tsne'

cat("\n<a href='#dimensionality-reduction'>");
<<visualizationIntegration>>
cat("</a> \n \n");

rm("useReduction");
```

#### UMAP
```{r visualizationIntegration_UMAP}
useReduction='umap'

cat("\n<a href='#dimensionality-reduction'>");
<<visualizationIntegration>>
cat("</a> \n \n");

rm("useReduction");
```

<!-- Ending the tabset section -->
### {.toc-ignore}



### Cell clustering from individual timepoint analysis {.tabset .tabset-fade}
#### Juvenile
```{r cellClustering_Projection_juvenile}
<<juvenile_clustering_umap>>
```

```{r cellClustering_Projection_juvenile_Splitted, fig.dim = c(3,3), out.width='25%'}
<<juvenile_clustering_umap_splitted>>

```

#### Adult
```{r cellClustering_Projection_adult}
<<adult_clustering_umap>>
```

```{r cellClustering_Projection_adult_Splitted, fig.dim = c(3,3), out.width='25%'}
<<adult_clustering_umap_splitted>>

```

#### Aged
```{r cellClustering_Projection_aged}
<<aged_clustering_umap>>
```

```{r cellClustering_Projection_aged_Splitted, fig.dim = c(3,3), out.width='25%'}
<<aged_clustering_umap_splitted>>

```

<!-- Ending the tabset section -->
### {.toc-ignore}




### Cell clustering {.tabset .tabset-fade}
```{r cellClusteringProcessing}
<<cellClustering>>
```


#### tSNE
```{r visualizationClustering_tSNE}
useReduction='tsne'

cat("\n<a href='#dimensionality-reduction'>");
<<cellClustering_visualization>>
cat("</a> \n \n");

rm("useReduction");
```

#### UMAP
```{r visualizationClustering_UMAP}
useReduction='umap'

cat("\n<a href='#dimensionality-reduction'>");
<<cellClustering_visualization>>
cat("</a> \n \n");

rm("useReduction");
```

<!-- Ending the tabset section -->
### {.toc-ignore}



### Identify conserved marker genes
#### Marker genes list
```{r identifyConservedMarkerGenes}
<<heterogeneity_conservedmarkerGenes>>
<<heterogeneity_conservedmarkerGenes_table>>
<<heterogeneity_conservedmarkerGenes_heatmap>>
<<heterogeneity_markerGenes>>
<<heterogeneity_markerGenes_table>>
<<heterogeneity_markerGenes_heatmap>>
```
<br /><br />


### Monitored genes
#### Monitored genes heatmap
```{r monitoredDotPlot}
<<heterogeneity_monitoredGenes_dotplot>>
```
<br /><br />

#### Monitored genes expression
```{r monitoredFeaturePlot, out.width='50%'}
<<heterogeneity_monitoredGenes_tsne_umap>>
```
<br /><br />

#### Monitored genes expression
```{r monitoredViolinPlot, out.width='50%'}
<<heterogeneity_monitoredGenes_vlnplot>>
```
<br /><br />

#### Monitored genes - density plot
```{r monitoredDensityPlot, out.width='50%'}
<<heterogeneity_monitoredGenes_densityplot>>
```
<br /><br />


### Exporting results
```{r saveSeuratObject}
<<save_data>>
<<extract_Lyve1_pos_cluster>>
```
<br /><br />

