---
output:
  html_document: 
    theme: cerulean
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r setKnitrOptions, include=FALSE}
knitr::opts_chunk$set(results='asis', echo = FALSE, error = TRUE, cache = FALSE, cache.comments = TRUE, fig.dim = c(7,7), autodep = TRUE)
```




```{r loadParamsAndDependencies, warning=FALSE, message=FALSE}
### Define working folder (contains R/Rmd file for current sample, parent contains global project files)
WORKING_DIR   = getwd();

### Load parameters
# Define an environment that will contain parameters
paramsEnv = new.env();

# Load file defining global (project) parameters in 'paramsEnv'
source( file.path( WORKING_DIR, "..", "globalParams.R"), local = paramsEnv);

# Load file defining local (sample) parameters eventually overriding global ones
sampleParamsFilePath = file.path( WORKING_DIR, "..", "sampleParams.R");
if(file.exists(sampleParamsFilePath)){
  source( sampleParamsFilePath, local = paramsEnv);
} else {
  warning("The file 'sampleParams.R' containing sample-specific parameters is missing. The analysis will use global (project) parameters only.");
}

# Load file defining analysis parameters
analysisParamsFilePath = file.path( WORKING_DIR, "analysisParams.R");
if(file.exists(analysisParamsFilePath)) {
  source( analysisParamsFilePath, local = paramsEnv);
} else {
  warning("The file 'analysisParamsFilePath.R' containing analysis-specific parameters is missing.");
}

# Assign loaded values to current environment (Not .GlobalEnv as it may not be the current one depending on where rendering is started from)
invisible( lapply( ls( paramsEnv, all.names = TRUE), function(x, envir)
  { 
    assign( x = x, 
            value = get( x, pos = paramsEnv), 
            pos = envir)
  }, 
  environment()));

### Load general Deps
cat("<BR>WORKING_DIR=", WORKING_DIR)
source( file.path( WORKING_DIR, "00_generalDeps.R"));

# Import chunks from source files
read_chunk( path=file.path( WORKING_DIR, "01_prepareData.R"));
read_chunk( path=file.path( WORKING_DIR, "02_variableGenes.R"));
read_chunk( path=file.path( WORKING_DIR, "03_cellHeterogeneity.R"));
read_chunk( path=file.path( WORKING_DIR, "04_macrophageSelection.R"));
```

---
title: `r paste( GLOBAL_DESCRIPTION, ANALYSIS_STEP_NAME, sep ="<BR>")`
---


```{r initFunctions}

### Create output directory
dir.create( PATH_ANALYSIS_OUTPUT, showWarnings = FALSE);


### Record starting time (for report and file names)
startTime = Sys.time();
startTimeReport   = format( startTime, "%Y %m %d %H:%M:%S")
startTimeFileName = format( startTime, "%Y_%m_%d_%H_%M_%S")


### Initialize utils

# Create a hash for all parameters
paramsHash = digest( paramsEnv, algo="murmur32");

# Create a 'safe' unique prefix for output files
# outputFilesPrefix = paste0( SCIENTIFIC_PROJECT_NAME, "_",     
#                             EXPERIMENT_PROJECT_NAME, "_"
#                             #startTimeFileName, "_",
#                             #paramsHash, "_"
#                             )

# Set seed for pseudo random number generator
set.seed(42);

# Create a hidden empty heatmap to prevent heatmapr interference with plotly functions (bug ?)
cat('<div style="display: none">');
htmltools::tagList(main_heatmap(matrix(0)));
cat('</div>');

# Create a hidden plotly to prevent bug when rendering multiple plots in report
cat('<div style="display: none">');
htmltools::tagList(plot_ly(x = rnorm(10), type = "histogram"));
cat('</div>');

# Create a hidden empty heatmap to prevent iheatmap interference with plotly functions
cat('<div style="display: none">');
htmltools::tagList(iheatmap(rnorm(10)));
cat('</div>');
```




## Primary analysis (QC)


### Run informations {.tabset .tabset-fade}

#### Run
<center>
```{r runInfo_run}
pander( c( 'Date'     = startTimeReport,
           'Hash'     = paramsHash));
```
</center>

#### Parameters
```{r runInfo_params}
#showSimpleDT( paramsEnv); # Does not handle lists or nested lists easily
listParams = as.list( paramsEnv);
pander(listParams[order(names(listParams))]); # Show params in alphabetical order
```

#### System
<center>
```{r runInfo_sysInfo}
showSimpleDT( Sys.info())
```
</center>

#### R
```{r runInfo_r}
pander( sessionInfo());
```

<!-- Ending the tabset section -->
### {.toc-ignore} 
<br /><br />

### Load data from quality control step
```{r dataLoadFilterNormalize}
<<loadData>>
<<normalizeData>>
```
<br /><br />


### Highly varibale genes
```{r findHVG_rep1}
<<findVariableGenes_seuratMethod>>
<<findVariableGenes_summaryPlot>>
<<findVariableGenes_summaryTable>>
```
<br /><br />


### Data integration {.tabset .tabset-fade}
This part shows the results of the combination of the data from the above replicate.
```{r dataIntegrationProcessing}
<<pca_tsne_umap>>
```


#### tSNE
```{r visualizationIntegration_tSNE}
useReduction='tsne'

cat("\n<a href='#dimensionality-reduction'>");
<<visualizationMergedData>>
cat("</a> \n \n");

rm("useReduction");
```

#### UMAP
```{r visualizationIntegration_UMAP}
useReduction='umap'

cat("\n<a href='#dimensionality-reduction'>");
<<visualizationMergedData>>
cat("</a> \n \n");

rm("useReduction");
```

<!-- Ending the tabset section -->
### {.toc-ignore}



### Cell clustering {.tabset .tabset-fade}
```{r cellClusteringProcessing}
<<cellClustering>>
```


#### tSNE
```{r visualizationClustering_tSNE}
useReduction='tsne'

cat("\n<a href='#dimensionality-reduction'>");
<<cellClustering_visualization>>
cat("</a> \n \n");

rm("useReduction");
```

#### UMAP
```{r visualizationClustering_UMAP}
useReduction='umap'

cat("\n<a href='#dimensionality-reduction'>");
<<cellClustering_visualization>>
cat("</a> \n \n");

rm("useReduction");
```

<!-- Ending the tabset section -->
### {.toc-ignore}


### Marker genes
```{r heterogeneity_markerGenes}
<<heterogeneity_markerGenes>>
```

#### Marker genes list
```{r heterogeneity_markerGenes_table}
<<heterogeneity_markerGenes_table>>
```
<br />

#### Marker genes heatmap

Expression heatmap of top marker genes sorted by Seurat clusters:

```{r heterogeneity_markerGenes_heatmap}
<<heterogeneity_markerGenes_heatmap>>
```

#### Marker genes expression on t-SNE/UMAP {.tabset .tabset-fade}

##### UMAP {.tabset .tabset-fade}
```{r heterogeneity_markerGenes_expression_tSNE, fig.dim = c(3,3), out.width='25%'}
useReduction='umap';

cat("\n<a href='#dimensionality-reduction'>");
<<cellClustering_visualization>>
cat("</a> \n \n");

<<heterogeneity_markerGenes_expression>>

rm("useReduction");
```

<!-- Ending the tabset section -->
##### {.toc-ignore}

##### tSNE {.tabset .tabset-fade}
```{r heterogeneity_markerGenes_expression_UMAP, fig.dim = c(3,3), out.width='25%'}
useReduction='tsne';

cat("\n<a href='#dimensionality-reduction'>");
<<cellClustering_visualization>>
cat("</a> \n \n");

<<heterogeneity_markerGenes_expression>>

rm("useReduction");
```

<!-- Ending the tabset section -->
##### {.toc-ignore}


<!-- Ending the tabset section -->
#### {.toc-ignore}
<br /><br />


#### Marker genes expression on violin plots

Here we show the detail of expression of all genes identified as marker for at least one cluster
as violin plot per cluster (with jitter point). Numbers above each violin plot is (i) top: the number
of cells with a positive expression of the gene in the cluster (ii) bottom : the number of cells
with a null expression of the gene in the cluster.

```{r heterogeneity_markerGenes_expression_violin, fig.keep = "all", fig.dim = c(3,3), out.width='25%'}
<<heterogeneity_markerGenes_expression_violinplot>>
```


### Monitored genes
```{r heterogeneity_monitoredGenes}
<<heterogeneity_monitoredGenes>>
```

#### Monitored genes heatmap

Expression heatmap of monitored genes sorted by Seurat clusters:

```{r heterogeneity_monitoredGenes_heatmap}
<<heterogeneity_monitoredGenes_heatmap>>
```

#### Monitored genes expression {.tabset .tabset-fade}

##### UMAP {.tabset .tabset-fade}
```{r heterogeneity_monitoredGenes_expression_tSNE, fig.dim = c(3,3), out.width='25%'}
useReduction='umap'

cat("\n<a href='#dimensionality-reduction'>");
<<cellClustering_visualization>>
cat("</a> \n \n");

<<heterogeneity_monitoredGenes_expression>>

rm("useReduction");
```

<!-- Ending the tabset section -->
##### {.toc-ignore}

##### tSNE {.tabset .tabset-fade}
```{r heterogeneity_monitoredGenes_expression_UMAP, fig.dim = c(3,3), out.width='25%'}
useReduction='tsne'

cat("\n<a href='#dimensionality-reduction'>");
<<heterogeneity_dimReduc_thumbnail>>
cat("</a> \n \n");

<<heterogeneity_monitoredGenes_expression>>

rm("useReduction");
```

<!-- Ending the tabset section -->
##### {.toc-ignore}

<!-- Ending the tabset section -->
#### {.toc-ignore}
<br /><br />

#### Monitored genes expression on dotplot
```{r heterogeneity_monitoredGenes_expression_dotplot, out.width='50%', fig.dim=c(6,8)}
<<heterogeneity_monitoredGenes_expression_dotplot>>
```
<br /><br />

#### Monitored genes expression on violin plots

Here we show the detail of expression of all monitored genes
as violin plot per cluster (with jitter point). Numbers above each violin plot is (i) top: the number
of cells with a positive expression of the gene in the cluster (ii) bottom : the number of cells
with a null expression of the gene in the cluster.

```{r heterogeneity_monitoredGenes_expression_violin, fig.keep = "all", fig.dim = c(3,3), out.width='25%'}
<<heterogeneity_monitoredGenes_expression_violinplot>>
```

#### Extract Lyve1 positive cluster
```{r extract_Lyve1_positive_cluster}
<<extract_Lyve1_pos_cluster>>
<<save_data>>
```


### Exporting results to file

```{r exportResults, results='asis', echo=FALSE, warning=FALSE}

cat("<BR>Exporting expression data to file:<BR>")

# Exporting raw expression data (as produced by Cellranger)
raw_expression_file_path = file.path( PATH_ANALYSIS_OUTPUT, paste0( outputFilesPrefix, "QCfiltered_raw_expression_matrix.csv"))
cat("<BR>* Raw expression:<BR>", raw_expression_file_path)
write.table( as.data.frame( GetAssayData(object = sc10x.rna.seurat.ju.all, slot = "counts")), file = raw_expression_file_path, quote = FALSE, col.names = TRUE, row.names = TRUE, sep=",")

# Exporting normalized expression data
normalized_expression_file_path = file.path( PATH_ANALYSIS_OUTPUT, paste0( outputFilesPrefix, "QCfiltered_Normalized_expression_matrix.csv"))
cat("<BR>* Normalized expression:<BR>", normalized_expression_file_path)
write.table( as.data.frame( GetAssayData(object = sc10x.rna.seurat.ju.all, slot = "data")), file = normalized_expression_file_path, quote = FALSE, col.names = TRUE, row.names = TRUE, sep=",")

cat("<BR><BR>Exporting Rdata to files:<BR>")

# Exporting Seurat object
seurat_object_file_path = file.path( PATH_ANALYSIS_OUTPUT, paste0( outputFilesPrefix, "QCfiltered_sc10x.rna.seurat.ju.all.RDS"))

cat("<BR>* Seurat Object:<BR>", seurat_object_file_path)
saveRDS( object = sc10x.rna.seurat.ju.all, file = seurat_object_file_path)


seurat_object_subset_file_path = file.path( PATH_ANALYSIS_OUTPUT, paste0( outputFilesPrefix, "QCfiltered_subset_sc10x.rna.seurat.ju.all.RDS"))

cat("<BR>* Seurat Object:<BR>", seurat_object_subset_file_path)
saveRDS( object = sc10x.rna.seurat.ju.subset, file = seurat_object_subset_file_path)

cat("<BR><BR>Exporting dimensionality reduction mapping to files:<BR>")

# Exporting PCA map
PCA_map_file_path = file.path( PATH_ANALYSIS_OUTPUT, paste0( outputFilesPrefix, "QCfiltered_Normalized_PCAcoord.tsv"))
cat("<BR>* PCA map:<BR>", PCA_map_file_path)
write.table( Embeddings( sc10x.rna.seurat.ju.all, reduction = "pca"), file = PCA_map_file_path, quote = FALSE, col.names = TRUE, row.names = TRUE, sep=",")

# Exporting t-SNE map
tsne_map_file_path = file.path( PATH_ANALYSIS_OUTPUT, paste0( outputFilesPrefix, "QCfiltered_Normalized_Tsnecoord.tsv"))
cat("<BR>* t-SNE map:<BR>", tsne_map_file_path)
write.table( Embeddings( sc10x.rna.seurat.ju.all, reduction = "tsne"), file = tsne_map_file_path, quote = FALSE, col.names = TRUE, row.names = TRUE, sep=",")

# Exporting UMAP map
umap_map_file_path = file.path( PATH_ANALYSIS_OUTPUT, paste0( outputFilesPrefix, "QCfiltered_Normalized_Umapcoord.tsv"))
cat("<BR>* UMAP map:<BR>", umap_map_file_path)
write.table( Embeddings( sc10x.rna.seurat.ju.all, reduction = "umap"), file = umap_map_file_path, quote = FALSE, col.names = TRUE, row.names = TRUE, sep=",")

cat("<BR><BR>Exporting clustering to files:<BR>")
seurat_clusters_file_path = file.path( PATH_ANALYSIS_OUTPUT, paste0( outputFilesPrefix, "QCfiltered_Normalized_Seurat_clusters.tsv"))
cat("<BR>Seurat Clusters:<BR>", seurat_clusters_file_path)
Idents( sc10x.rna.seurat.ju.all) = "seurat_clusters"
write.table( data.frame( Idents( sc10x.rna.seurat.ju.all)), file = seurat_clusters_file_path, quote = FALSE, col.names = TRUE, row.names = TRUE, sep=",")

cat("<BR><BR>Exporting marker genes:<BR>")
gene_markers_file_path = file.path( PATH_ANALYSIS_OUTPUT, paste0( outputFilesPrefix, "QCfiltered_Normalized_Marker_Genes.csv"))
cat("<BR>Marker genes:<BR>", gene_markers_file_path)
write.table( markers, file = gene_markers_file_path, quote = FALSE, col.names = TRUE, row.names = TRUE, sep="\t")


```

